---
title: "RExercise#2_berganio"
author: "Laiza Joy Berganio"
date: "2026-09-30"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part A — Arithmetic Operations

```{r}
# TODO: Store 17 in a variable called a, and 5 in a variable called b
# Then print the results of: a + b, a - b, a * b, a / b

a <- 17
b <- 5

a+b
a-b
a*b
a/b
```

[1] 22

[1] 12

[1] 85

[1] 3.4

```{r}
# TODO: Using the same a and b, compute and print a %% b and a %/% b
# a %% b = remainder of a / b
# a %/% b = whole-number (integer) division

a%%b
a%/%b
```

[1] 2

[1] 3

```{r}
# TODO: Print 2 raised to the power of 10 using the ^ operator

2^10
```

[1] 1024

**Question #1:** Run 17 / 5, 17 %% 5, and 17 %/% 5 side by side. In your
own words, explain how these three results relate to each other
mathematically.

The '/' operator performs division, '%%' performs modulo which gets the
remainder of the values, and the '%/%' is the integer division operator
which divides the values but only takes the integer

**Question #2:** Predict the output of `10 %% 3` and `10 %/% 3` *before*
running the code. Were you correct? If not, explain what you
misunderstood.

The output of \> 10 %% 3 is 1 and 10 %/% is 3. These are correct.

------------------------------------------------------------------------

### Part B — Basic Data Types: typeof(), class(), and mode()

```{r}
x <- 42
y <- 42L        # the L suffix forces an integer
z <- "42"
w <- TRUE
v <- 3 + 2i     # complex number

# TODO: Use typeof() to check the type of each of x, y, z, w, v

typeof(x)
typeof(y)
typeof(z)
typeof(w)
typeof(v)
```

[1] "double"

[1] "integer"

[1] "character"

[1] "logical"

[1] "complex"

```{r}
# TODO: Use class() to check the class of each of x, y, z, w, v

class(x)
class(y)
class(z)
class(w)
class(v)
```

[1] "numeric"

[1] "integer"

[1] "character"

[1] "logical"

[1] "complex"

```{r}
# TODO: Use mode() to check the class of each of x, y, z, w, v

mode(x)
mode(y)
mode(z)
mode(w)
mode(v)
```

[1] "numeric"

[1] "numeric"

[1] "character"

[1] "logical"

[1] "complex"

**Question #3):** `x <- 42` and `y <- 42L` look almost identical, but
`typeof(x)` and `typeof(y)` give different answers. Explain the
difference in your own words.

`typeof()` and `typeof(y)` give different answers because in x, 42 is
treated as a decimal, while in y, 42 is treated as an integer due to the
'L' suffix. The 'L' suffix forces R to store the number as a whole
integer. If the stored value doesnt have an L, R stores the number as a
decimal.

**Question #4:** Is `typeof()` ever different from `class()` and from
`mode()` for the same object? Look at your results above and explain any
case where this happens (or, if you did not see one, explain what you
would expect for a `matrix` or `factor`).\
Yes. In x, its type is a "double", but its class and mode are "numeric".
This is because typeof() identifies the objects' storage type, class()
identifies the object class, and mode() identifies its general data
category.

------------------------------------------------------------------------

### Part C — Type Conversion

```{r}
num_string <- "123"

# TODO: Convert num_string to a numeric value using as.numeric()
# Then check its type with typeof(), class(), and mode()

as.numeric(num_string)

typeof(num_string)
class(num_string)
mode(num_string)
```

[1] "character"

[1] "character"

[1] "character"

```{r}
# TODO: Convert the number 0 and the number 1 to logical using as.logical()
# What do you get?

as.logical(1)
as.logical(0)
```

[1] TRUE

[1] FALSE

```{r}
bad_string <- "hello123"

# TODO: Try converting bad_string to numeric using as.numeric()
# Run this chunk and observe the warning message and the result

as.numeric(bad_string)
```

[1] NA

Warning message: NAs introduced by coercion

**Question #5:** When you ran the `conversion-fail` chunk, R produced
`NA` and a warning instead of stopping with an error.

- Why do you think R is designed to give a warning here rather than
  crashing the whole script?

  R gives a warning so that the rest of the program can be executed even
  if a value cannot be converted. It also prevents data loss by just
  giving a warning rather than crashing.

- What are the risks of this behavior if you don’t read your warnings?

  It can affect calculations and lead to incorrect outputs, which causes
  errors in the results.

**Question #6:** `as.logical(0)` and `as.logical(1)` convert to `FALSE`
and `TRUE`.

- What do you think `as.logical(2)` and as.logical(-1)\` will produce?

  I think both will produce `TRUE` because any nonzero number, whether
  positive or negative, is treated as `TRUE` in R.

- Test it, then explain the rule R seems to be following.

  ```{r}
  as.logical(2)
  as.logical(-1)
  ```

  R seems to be following a rule that a zero is converted to `FALSE`,
  while any number other than zero is converted to `TRUE`, whether the
  number is positive or negative.

------------------------------------------------------------------------

### Part D — Strings, paste(), and paste0()

```{r}
first_name <- "Juan"
last_name <- "Dela Cruz"

# TODO: Print first_name and last_name using print() and/or cat()
```

```{r}
cat(first_name,last_name)
```

[1] Juan Dela Cruz

```{r}
# TODO: Use paste() to combine first_name and last_name into a full name
# TODO: Use paste0() to do the same, and compare the two results
paste(first_name,last_name)
paste0(first_name,last_name)
```

[1] "Juan Dela Cruz" [1] "JuanDela Cruz"

```{r}
# TODO: Use paste() with the argument sep = "-" to join first_name and
# last_name with a dash instead of a space
paste(first_name,last_name, sep = "-")
```

[1] "Juan-Dela Cruz"

```{r}
age <- 21
# TODO: Use paste0() to build the sentence:
# "Juan Dela Cruz is 21 years old." by combining full_name, age, and text

paste0(first_name, " ",last_name, " is ", age, " years old.")
```

[1] "Juan Dela Cruz is 21 years old."

**Question #7:** `paste()` and `paste0()` produced different results for
the same two strings. Explain exactly why, referencing what each
function does by default.

`paste()` adds a space between strings by default, while `paste0()`
combines strings directly without adding a separator. This is why their
outputs are different even when given the same two strings.

**Question #8:** In the `strings-numbers` chunk, you combined a
character string with a numeric variable (`age`) using `paste0()`
without first converting `age` to a string.

- Did this cause an error? Explain what `paste0()` does internally to
  make this work, and how it relates to the type conversions you
  practiced in Part C.

  No, it did not cause an error because `paste0()` automatically
  converts the number into a character string before combining it with
  the text. This is similar to using `as.character()` to convert other
  data types into strings.

------------------------------------------------------------------------

### Part E — Critical Thinking Challenge

```{r}
# TODO: Predict, then test: what is the result of "5" + 5 in R?
# Run it and read the error message carefully.

# "5" + 5
```

Error in "5" + 5 : non-numeric argument to binary operator

```{r}
# TODO: Now fix the code above so that "5" + 5 correctly evaluates to 10
# using an appropriate conversion function.
as.numeric("5")+5
```

[1] 10

**Question #9:** Some programming languages automatically convert
`"5" + 5` into `10` without an error. R does not.

- Based on what you’ve learned about R’s data types in this lab, why do
  you think R requires an explicit conversion instead of guessing what
  the programmer meant?

*(Your answer here)*

**Question #10:** Suppose a classmate writes a script that reads a
column of “numbers” from a text file. The column looks numeric, but when
they try `sum()` on it, R throws an error. Based on everything in this
lab (types, `typeof()`/`class()`/`mode()`, conversion, warnings), write
a short paragraph explaining what likely went wrong and the steps you’d
take to diagnose and fix it.

*(Your answer here)*
